---
title: "Function-b1"
author: "James Forward"
date: "2023-11-01"
output: md_document
Roxygen: list(markdown = TRUE)
---

```{r setup, include=FALSE}
library(datateachr)
library(tidyverse)
library(palmerpenguins)
library(dplyr)
library(testthat)
```
### Exercise 1 + 2: Make a Function + Document your Function

- I made a function that allows me to take the data frame 'penguins' and create a new data frame with the mean values of specified measurements for each species. 

```{r}

#' Summarize Measurements by Species
#'
#' This function returns a summarized data frame with mean measurements grouped by species. 
#' The user can specify which variables to summarize.
#'
#' @param data A data frame containing at least a "species" column and columns specified in `summary_vars`.
#' @param summary_vars A character vector indicating which columns in `data` to summarize. 
#'                     Default is c("bill_length_mm", "flipper_length_mm", "body_mass_g").
#' @param ... Additional arguments passed on to the `mean` function, such as `na.rm`.
#' 
#' @return A data frame with one row for each species and a column for the mean of each `summary_vars`.
#'
#' @export

summarize_by_species <- function(data, summary_vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g"), ...) {
  # Error trap to ensure that the provided data is a data frame
  if (!is.data.frame(data)) {
    stop("Input data must be a data frame.")
  }
  
  if (!all(summary_vars %in% names(data))) {
    stop("Not all specified summary variables are present in the data frame.")
  }

 # Group by species and compute the mean for specified variables
  summarized_data <- data %>%
    group_by(species) %>%
    summarise(across(all_of(summary_vars), mean, na.rm = TRUE, .names = "mean_{.col}"), ...)

  return(summarized_data)
}

```

### Exercise 3: Examples

#### Example 1: Default function to see if it works


```{r}

#Using the penguins data, I will use the function 
result <- summarize_by_species(penguins)
print(result)

```

#### Example 2: Fake Numeric Variable

```{r}

# Here I created a variable that is categorical but is actually represented with numbers
penguins$fake_numeric <- as.factor(sample(1:5, nrow(penguins), replace = TRUE))

# Trying to get the mean of this fake numeric variable using our function
result_error <- summarize_by_species(penguins, summary_vars = c("bill_length_mm", "flipper_length_mm", "body_mass_g", "fake_numeric"))

print(result_error)
```

### Exercise 4: Test the Function
```{r}

test_that("summarize_by_species behaves correctly", {

# Test 1: Vector that has NAs
  # Assuming 'body_mass_g' contains NAs
  result1 <- summarize_by_species(penguins, summary_vars = "body_mass_g")
  expect_equal(dim(result1), c(3, 2))
  expect_true("species" %in% colnames(result1))
  expect_true("mean_body_mass_g" %in% colnames(result1))
  
  # Test 2: Vector of a different type (using a character vector for numeric_category which should generate a warning or error)
  penguins$char_col <- sample(letters, nrow(penguins), replace = TRUE)
  expect_warning(summarize_by_species(penguins, summary_vars = "char_col"))
  
  # Test 3: Vector of length 0 (like numeric(0))
  expect_error(summarize_by_species(penguins, summary_vars = "nonexistent_variable"))
  
})

